---
title: "K-12 Computer Science Course Analysis (2020–2022)"
author: "Omodolapo Ojo, PhD"
date: !r Sys.Date()
output:
  html_document:
    toc: true
    toc_depth: 2
  word_document:
    toc: true
  pdf_document:
    toc: true
---

## Project Overview
This project explores the availability of **Computer Science (CS) classes** in K-12 schools across the United States using the **Civil Rights Data Collection (CRDC)** and **Common Core of Data (CCD)** data sets for 2020–21 and 2021–22 school years.

The main goal is to **clean, merge, and prepare the data** for analysis, ensuring consistent variable names, proper handling of missing or inconsistent values, and meaningful recoding of categorical variables such as **Title I status** and **Virtual instruction mode**.

Overall, the data cleaning process shows a solid foundation in clean data engineering

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(dplyr)
library(tidyr) #to prepare the identifier variables
library(ggplot2)
library(readr)
library(stringr)
```

## Introduction

This notebook demonstrates the cleaning, merging, and preparation of K-12 CS datasets for analysis. We use CRDC and CCD data for 2020–21 and 2021–22. 
Civil Rights Data Collection (CRDC), public access portal is https://ocrdata.ed.gov/; 
Common Core of Data (CCD) https://nces.ed.gov/ccd/, then click the Direct data files page, CCD Data Files

## Load Datasets
```{r, echo = FALSE}
#The main dataset with info on CS enrollment by gender and race
data_crdc_2021<-read.csv("data/ComputerScience_2021_2022.csv")

data_crdc_2020 <-read.csv("data/ComputerScience_2020_2021.csv")

#The sub datasets with info on status of schools, based on title I, virtual instructions
data_ccd_2020 <-read.csv("data/ccd_sch_129_2021_w_1a_080621__school_2020__2021.csv")

data_ccd_2021 <-read.csv("data/ccd_sch_129_2122_w_1a_071722_school_characteristics.csv")

#Another sub datasets primarily added because it has info on school size and grades
data_ccd_membership_2021<-read.csv("data/ccd_SCH_052_2122_l_1a_071722.csv")

data_ccd_membership_2020<-read.csv("data/ccd_SCH_052_2021_l_1a_080621.csv")

```

## Data Cleaning
1. **Variable Selection and Harmonization**:  
   - Selected only the needed variables from the sub-datasets
   - Renamed all variables to consistent, human-readable names.  
   - Standardized identifiers (`Combined_Key` / `NCESSCH`) to 12-character strings for safe joins.

```{r, echo = FALSE}
#Select from each of the sub-dataset only the relevant variables
data_ccd_2020 <- data_ccd_2020 %>%
  select(SCHOOL_YEAR, NCESSCH, TITLEI_STATUS_TEXT, VIRTUAL_TEXT)

data_ccd_2021 <- data_ccd_2021 %>%
  select(SCHOOL_YEAR, NCESSCH, TITLEI_STATUS_TEXT, VIRTUAL_TEXT)

data_ccd_membership_2020 <- data_ccd_membership_2020 %>%
  select(NCESSCH, GRADE, STUDENT_COUNT)

data_ccd_membership_2021 <- data_ccd_membership_2021 %>%
  select(NCESSCH, GRADE, STUDENT_COUNT)

#--Renamed all variables to consistent, human-readable names.
#Create a naming map
#Use human-readable, meaningful names. Keep them identical across datasets
#Remember to do LEP → EL harmonization for the 2020 CRDC data. 
name_map <- c(LEA_STATE = "State_Code",
  LEA_STATE_NAME = "State_Name",
  LEAID = "District_ID",
  LEA_NAME = "District_Name",
  SCHID = "School_ID",
  SCH_NAME = "School_Name",
  COMBOKEY = "Combined_Key",
  JJ = "Juvenile_Justice_School",
  SCH_COMPCLASSES_CSCI = "CS_Classes_Offered",
  SCH_COMPENR_CSCI_HI_M = "Hispanic_Male",
  SCH_COMPENR_CSCI_HI_F = "Hispanic_Female",
  SCH_COMPENR_CSCI_AM_M = "AmericanIndian_Male",
  SCH_COMPENR_CSCI_AM_F = "AmericanIndian_Female",
  SCH_COMPENR_CSCI_AS_M = "Asian_Male",
  SCH_COMPENR_CSCI_AS_F = "Asian_Female",
  SCH_COMPENR_CSCI_HP_M = "PacificIslander_Male",
  SCH_COMPENR_CSCI_HP_F = "PacificIslander_Female",
  SCH_COMPENR_CSCI_BL_M = "Black_Male",
  SCH_COMPENR_CSCI_BL_F = "Black_Female",
  SCH_COMPENR_CSCI_WH_M = "White_Male",
  SCH_COMPENR_CSCI_WH_F = "White_Female",
  SCH_COMPENR_CSCI_TR_M = "TwoOrMore_Male",
  SCH_COMPENR_CSCI_TR_F = "TwoOrMore_Female",
  
  TOT_COMPENR_CSCI_M = "Total_Male",
  TOT_COMPENR_CSCI_F = "Total_Female",
  
  SCH_COMPENR_CSCI_EL_M = "EnglishLearner_Male",
  SCH_COMPENR_CSCI_EL_F = "EnglishLearner_Female",
  
  SCH_COMPENR_CSCI_IDEA_M = "Disability_Male",
  SCH_COMPENR_CSCI_IDEA_F = "Disability_Female"
)

#Apply re-labeling to data_crdc_2021
name_map_rev <- setNames(names(name_map), name_map)
data_crdc_2021_renamed <- data_crdc_2021 %>%
  rename(any_of(name_map_rev))
str(data_crdc_2021)

#Ensure LEP→EL harmonization first, before relabeling data_crdc_2020
data_crdc_2020 <- data_crdc_2020 %>%
  rename(
    SCH_COMPENR_CSCI_EL_M = SCH_COMPENR_CSCI_LEP_M,
    SCH_COMPENR_CSCI_EL_F = SCH_COMPENR_CSCI_LEP_F
  ) %>%
  rename(any_of(name_map))

#Apply the common re-labeling to data_crdc_2020
name_map_rev <- setNames(names(name_map), name_map)
data_crdc_2020_renamed <- data_crdc_2020 %>%
  rename(any_of(name_map_rev))
str(data_crdc_2020_renamed)

#--Standardized identifiers (`Combined_Key` / `NCESSCH`) to 12-character strings for safe joins
#Verified uniqueness of keys.
#First, ensure the identifier or merge variables in both datasets look alike
data_crdc_2021_renamed <- data_crdc_2021_renamed %>%
  mutate(Combined_Key = str_pad(Combined_Key, 12, pad = "0"))
 
data_ccd_2021 <- data_ccd_2021 %>%
  mutate(NCESSCH = str_pad(NCESSCH, 12, pad = "0"))

data_ccd_membership_2021 <- data_ccd_membership_2021 %>%
  mutate(
    NCESSCH = sprintf("%012.0f", NCESSCH)
  )

#Repeat the same process for the other datasets
# First, ensure the key or identifier or merge variables look alike
data_crdc_2020_renamed <- data_crdc_2020_renamed %>%
  mutate(Combined_Key = str_pad(Combined_Key, 12, pad = "0"))

#NCESSCH in 2020 CCD data is numeric (double) while Combined_Key is a character. But left_join() requires both columns to be the same type
data_ccd_2020 <- data_ccd_2020 %>%
  mutate(NCESSCH = as.character(NCESSCH))

data_ccd_2020 <- data_ccd_2020 %>%
  mutate(NCESSCH = str_pad(NCESSCH, 12, pad = "0"))

## Re-encode NCESSCH as a zero-padded character string. It should be 12 digits like the already existing data (CCD and CDRC)
data_ccd_membership_2020 <- data_ccd_membership_2020 %>%
  mutate(
    NCESSCH = sprintf("%012.0f", NCESSCH)
  )

```

2. **Merging CRDC & CCD datasets**:  
   - Verified uniqueness of keys.  
   - Left-joined to combine datasets, CRDC (school-level outcomes), CCD (school characteristics), CCD membership (enrollment totals)

```{r, echo = FALSE}
#--Check uniqueness of keys before joining
# Left table (should usually be unique)
data_crdc_2021_renamed %>%
  count(Combined_Key) %>%
  filter(n > 1)

# Right table (must be unique for safe join)
data_ccd_2021 %>%
  count(NCESSCH) %>%
  filter(n > 1)

#List all the distinct title I status
data_ccd_2021 %>%
  distinct(TITLEI_STATUS_TEXT)

#Title I grouping
data_ccd_2021 <- data_ccd_2021 %>%
  mutate(
 title1_group = case_when(
      TITLEI_STATUS_TEXT %in% c(
        "Title I schoolwide school",
        "Title I schoolwide eligible school-No program"
      ) ~ "Title I - Schoolwide",

      TITLEI_STATUS_TEXT %in% c(
        "Title I targeted assistance school",
        "Title I targeted assistance eligible school-No program",
        "Title I schoolwide eligible-Title I targeted assistance program"
      ) ~ "Title I - Targeted Assistance",

      TITLEI_STATUS_TEXT == "Not a Title I school" ~ "Not Title I",

      TITLEI_STATUS_TEXT %in% c("Not reported", "Missing") ~ NA_character_,

      TRUE ~ NA_character_
    )
  )


data_ccd_2021 <- data_ccd_2021 %>%
  group_by(NCESSCH) %>%
  mutate(
    has_title1_info = any(!is.na(TITLEI_STATUS_TEXT)),
    
    TitleI_Schoolwide = if_else(
      has_title1_info,
      as.integer(any(TITLEI_STATUS_TEXT %in% c(
        "Title I schoolwide school",
        "Title I schoolwide eligible school-No program"
      ))),
      NA_integer_
    ),
    
    TitleI_Targeted = if_else(
      has_title1_info,
      as.integer(any(TITLEI_STATUS_TEXT %in% c(
        "Title I targeted assistance school",
        "Title I targeted assistance eligible school-No program",
        "Title I schoolwide eligible-Title I targeted assistance program"
      ))),
      NA_integer_
    ),
    
    Not_TitleI = if_else(
      has_title1_info,
      as.integer(any(TITLEI_STATUS_TEXT == "Not a Title I school")),
      NA_integer_
    )
  ) %>%
  ungroup() %>%
  select(-has_title1_info)
    
##Because the CCD membership data has multiple rows per school, then we need to first sort the variables before ensuring uniqueness
#Listed all grade values exactly as they appear in the data for decision how to group them 

#Listed all the distinct grades
data_ccd_membership_2021 %>%
  distinct(GRADE)

#Grouped the grades accordingly
data_ccd_membership_2021 <- data_ccd_membership_2021 %>%
  mutate(
    grade_group = case_when(
      GRADE %in% c(
        "Pre-Kindergarten", "Kindergarten",
        "Grade 1", "Grade 2", "Grade 3", "Grade 4", "Grade 5"
      ) ~ "Elementary",

      GRADE %in% c(
        "Grade 6", "Grade 7", "Grade 8"
      ) ~ "Middle",

      GRADE %in% c(
        "Grade 9", "Grade 10", "Grade 11", "Grade 12"
      ) ~ "High",

      GRADE %in% c(
        "Not Specified", "No Category Codes", "Ungraded", "Adult Education"
      ) ~ NA_character_,

      TRUE ~ NA_character_
    )
  )

#Now collapsed to one row per school, by summarizing the student counts as school enrollment, and producing three columns (Elementary, Middle, High) with 0 / 1 / NA.
data_ccd_membership_2021 <- data_ccd_membership_2021 %>%
  group_by(NCESSCH) %>%
  mutate(
    school_enrollment = sum(STUDENT_COUNT, na.rm = TRUE),
    
    has_any_grade_info = any(!is.na(grade_group)),

    Elementary = if_else(
      has_any_grade_info,
      as.integer(any(grade_group == "Elementary")),
      NA_integer_
    ),

    Middle = if_else(
      has_any_grade_info,
      as.integer(any(grade_group == "Middle")),
      NA_integer_
    ),

    High = if_else(
      has_any_grade_info,
      as.integer(any(grade_group == "High")),
      NA_integer_
    )
  ) %>%
  ungroup() %>%
  select(-has_any_grade_info)

#Ensure the uniqueness of NCESSCH as well for the CCD membership file
data_ccd_membership_2021 %>%
  count(NCESSCH) %>%
  filter(n > 1)

#Repeat the same process for other datasets
# Left table (should usually be unique)
data_crdc_2020_renamed %>%
  count(Combined_Key) %>%
  filter(n > 1)

# Right table (must be unique for safe join)
data_ccd_2020 %>%
  count(NCESSCH) %>%
  filter(n > 1)

#List all the distinct title I status
data_ccd_2020 %>%
  distinct(TITLEI_STATUS_TEXT)

#Title I grouping
data_ccd_2020 <- data_ccd_2020 %>%
  mutate(
 title1_group = case_when(
      TITLEI_STATUS_TEXT %in% c(
        "Title I schoolwide school",
        "Title I schoolwide eligible school-No program"
      ) ~ "Title I - Schoolwide",

      TITLEI_STATUS_TEXT %in% c(
        "Title I targeted assistance school",
        "Title I targeted assistance eligible school-No program",
        "Title I schoolwide eligible-Title I targeted assistance program"
      ) ~ "Title I - Targeted Assistance",

      TITLEI_STATUS_TEXT == "Not a Title I school" ~ "Not Title I",

      TITLEI_STATUS_TEXT %in% c("Not reported", "Missing") ~ NA_character_,

      TRUE ~ NA_character_
    )
  )

data_ccd_2020 <- data_ccd_2020 %>%
  group_by(NCESSCH) %>%
  mutate(
    has_title1_info = any(!is.na(TITLEI_STATUS_TEXT)),
    
    TitleI_Schoolwide = if_else(
      has_title1_info,
      as.integer(any(TITLEI_STATUS_TEXT %in% c(
        "Title I schoolwide school",
        "Title I schoolwide eligible school-No program"
      ))),
      NA_integer_
    ),
    
    TitleI_Targeted = if_else(
      has_title1_info,
      as.integer(any(TITLEI_STATUS_TEXT %in% c(
        "Title I targeted assistance school",
        "Title I targeted assistance eligible school-No program",
        "Title I schoolwide eligible-Title I targeted assistance program"
      ))),
      NA_integer_
    ),
    
    Not_TitleI = if_else(
      has_title1_info,
      as.integer(any(TITLEI_STATUS_TEXT == "Not a Title I school")),
      NA_integer_
    )
  ) %>%
  ungroup() %>%
  select(-has_title1_info)    

##Because the CCD membership data has multiple rows per school, then we need to first sort the variables before ensuring uniqueness
#Listed all grade values exactly as they appear in the data for decision how to group them 
#Listed all the distinct grades
data_ccd_membership_2020 %>%
  distinct(GRADE)

#Grouped the grades accordingly
data_ccd_membership_2020 <- data_ccd_membership_2020 %>%
  mutate(
    grade_group = case_when(
      GRADE %in% c(
        "Pre-Kindergarten", "Kindergarten",
        "Grade 1", "Grade 2", "Grade 3", "Grade 4", "Grade 5"
      ) ~ "Elementary",

      GRADE %in% c(
        "Grade 6", "Grade 7", "Grade 8"
      ) ~ "Middle",

      GRADE %in% c(
        "Grade 9", "Grade 10", "Grade 11", "Grade 12"
      ) ~ "High",

      GRADE %in% c(
        "Not Specified", "No Category Codes", "Ungraded", "Adult Education"
      ) ~ NA_character_,

      TRUE ~ NA_character_
    )
  )

#Now collapsed to one row per school, by summarizing the student counts as school enrollment, and producing three columns (Elementary, Middle, High) with 0 / 1 / NA.
data_ccd_membership_2020 <- data_ccd_membership_2020 %>%
  group_by(NCESSCH) %>%
  mutate(
    school_enrollment = sum(STUDENT_COUNT, na.rm = TRUE),
    
    has_any_grade_info = any(!is.na(grade_group)),

    Elementary = if_else(
      has_any_grade_info,
      as.integer(any(grade_group == "Elementary")),
      NA_integer_
    ),

    Middle = if_else(
      has_any_grade_info,
      as.integer(any(grade_group == "Middle")),
      NA_integer_
    ),

    High = if_else(
      has_any_grade_info,
      as.integer(any(grade_group == "High")),
      NA_integer_
    )
  ) %>%
  ungroup() %>%
  select(-has_any_grade_info)

#Ensure the uniqueness of NCESSCH
data_ccd_membership_2020 %>%
  count(NCESSCH) %>%
  filter(n > 1)

#--Left-joined to combine school-level characteristics and CS enrollment data.
#Now merge the first pair of datasets
data_crdc_ccd_2021 <- data_crdc_2021_renamed %>%
  left_join(data_ccd_2021, by = c("Combined_Key" = "NCESSCH")) %>%
  left_join(data_ccd_membership_2021, by = c("Combined_Key" = "NCESSCH"))

#Now merge the second pair of datasets
data_crdc_ccd_2020 <- data_crdc_2020_renamed %>%
  left_join(data_ccd_2020, by = c("Combined_Key" = "NCESSCH")) %>%
  left_join(data_ccd_membership_2020, by = c("Combined_Key" = "NCESSCH"))

# Be sure the combined datasets did not introduce duplicates
nrow(data_crdc_ccd_2021) == nrow(data_crdc_2021_renamed)

nrow(data_crdc_ccd_2020) == nrow(data_crdc_2020_renamed)

#Now the Next Step is to check whether the two data-sets match
setdiff(names(data_crdc_ccd_2021), names(data_crdc_ccd_2020)) #Variable in data_crdc_ccd_2021 but not in data_crdc_ccd_2020

setdiff(names(data_crdc_ccd_2020), names(data_crdc_ccd_2021)) #Variable in data_crdc_ccd_2020 but not in data_crdc_ccd_2021

# Since data_crdc_ccd_2020 has one extra column that data_crdc_ccd_2021 does not, the clean way is to keep only the columns common to both. The immediate code does the following: keeps only columns that exist in both data frames; preserves column order as they appear in data_crdc_ccd_2021 and works even if there are multiple extra columns
data_crdc_ccd_2020_reduced <- data_crdc_ccd_2020[, intersect(names(data_crdc_ccd_2020), names(data_crdc_ccd_2021))]

#Now you can combine the two datasets to form a long format data
combined_data <- bind_rows(data_crdc_ccd_2021, data_crdc_ccd_2020_reduced)
str(combined_data)

```

3. **Handling Missing & Invalid Values**:  
   - Replaced negative counts with `NA`.  
 
```{r, echo = FALSE}
# Replace all negative values with NA in the data-set
combined_data_clean <- combined_data %>%
  mutate(across(
    where(is.numeric),       # Only numeric columns
    ~ ifelse(.x < 0, NA, .x) # Replace negative numbers with NA
  ))
str(combined_data_clean)

```
   
4. **Derived Variables**:  
   - Total CS enrollment by gender and race.  
   - Race and gender proportions.  
   - Title I and virtual instruction groups.
  - Dropped rows with inconsistencies
```{r, echo = FALSE}
#Create race-specific totals, as well as total enrollment (male + female). It's okay to worry about multi-collinearity here: total enrollment is a linear combination of sex and race counts. You cannot include all of them in their raw-count form in the same model.

#The key idea is re-parameterization: keep one scale variable (total enrollment) and express everything else as relative composition (proportions or ratios) with a baseline (reference) category.

combined_data_clean1 <- combined_data_clean %>%
  mutate(
    # Totals
    Total_CS_Enrolment  = rowSums(cbind(Total_Male, Total_Female), na.rm = TRUE),
    White_CS        = rowSums(cbind(White_Male, White_Female), na.rm = TRUE),
    Black_CS        = rowSums(cbind(Black_Male, Black_Female), na.rm = TRUE),
    Hispanic_CS     = rowSums(cbind(Hispanic_Male, Hispanic_Female), na.rm = TRUE),
    Asian_CS        = rowSums(cbind(Asian_Male, Asian_Female), na.rm = TRUE),
    AmericanIndian_CS = rowSums(cbind(AmericanIndian_Male, AmericanIndian_Female), na.rm = TRUE),
    Multiracial_CS  = rowSums(cbind(TwoOrMore_Male, TwoOrMore_Female), na.rm = TRUE)
  ) 

#Protect against division by zero
combined_data_clean1 <- combined_data_clean1 %>%
  mutate(
    Total_CS_Enrolment = ifelse(Total_CS_Enrolment == 0, NA, Total_CS_Enrolment)
  )

#The base category for male and female gender proportion is the total CS enrollment
combined_data_clean1 <- combined_data_clean1 %>%
  mutate(
    Prop_Male_CS = Total_Male / Total_CS_Enrolment,
    Prop_Female_CS = Total_Female / Total_CS_Enrolment,
)

#Quick sanity check
combined_data_clean1 %>%
  select(starts_with("Prop_")) %>%
  summary()

#Take a diagnostic step. You can identify exactly which observations (and which places) have race totals exceeding total enrollment, and by how much. 
#Create a race-sum variable and compare it to total enrollment
combined_data_clean1 <- combined_data_clean1 %>%
  mutate(
    Race_Total_CS =
      White_CS + Black_CS + Hispanic_CS + Asian_CS +
      AmericanIndian_CS + Multiracial_CS,
    
    Race_Exceeds_Total = Race_Total_CS > Total_CS_Enrolment
  )

#How many observations are affected?
table(combined_data_clean1$Race_Exceeds_Total, useNA = "ifany")

#Drop the 19 inconsistent rows (recommended)
#They carry almost no information and can distort coefficients.
combined_data_clean1 <- combined_data_clean1 %>%
  filter(!Race_Exceeds_Total | is.na(Race_Exceeds_Total))

#Recomputing proportions after filtering is very important. However, total enrollment is not a safe denominator everywhere, because some rows have sex totals only. some rows have race totals only, some have both, some have neither. Hence, there is no single denominator that is valid for all rows.
combined_data_clean1 <- combined_data_clean1 %>%
  mutate(
  Total_Race_CS =
     White_CS + Black_CS + Hispanic_CS + Asian_CS +
    AmericanIndian_CS + Multiracial_CS,
  
  Total_Race_CS = ifelse(Total_Race_CS == 0, NA, Total_Race_CS),
   
   Prop_White_CS = White_CS / Total_Race_CS,
   Prop_Hispanic_CS = Hispanic_CS / Total_Race_CS,
   Prop_Asian_CS = Asian_CS / Total_Race_CS,
   Prop_AmericanIndian_CS = AmericanIndian_CS / Total_Race_CS,
   Prop_Multiracial_CS = Multiracial_CS / Total_Race_CS
   )

#Sanity check to ensure all race proportions ≤ 1
#No Inf, and NA’s only from missing totals
combined_data_clean1 %>%
  select(starts_with("Prop_")) %>%
  summary()

#Now, all rows where the sum of race-based CS enrollment does not equal the reported total CS enrollment are flagged, and then counted to know how often that happens.
combined_data_clean_Race_Checked <- combined_data_clean1 %>% mutate( Race_Not_Equal_Total = !is.na(Total_CS_Enrolment) & !is.na(Race_Total_CS) & Race_Total_CS != Total_CS_Enrolment )

table(combined_data_clean_Race_Checked$Race_Not_Equal_Total, useNA = "ifany")

#Even after removing the 19 rows (i.e., rows where race total exceeds total enrollment), we still found that race totals ≠ total enrollment for 2,760 rows. Meanwhile, the rows remain usable. Only that there's an assurance that total enrollment is not a universal denominator. Therefore, race-based denominators for race are statistically correct

#Final sanity check to there are no cases of schools where there no CS courses, yet there are reported cases of CS enrollment. First, identify logically inconsistent schools. There were no such cases
# Flag rows where there are no CS classes offered, yet total enrollment > 0
combined_data_clean_Courses_Checked <- combined_data_clean1 %>%
  mutate(
    Enrollment_Without_CS_Courses = 
      CS_Classes_Offered == 0 & Total_CS_Enrolment > 0
  )

# Report counts. There were no cases of such
table(combined_data_clean_Courses_Checked$Enrollment_Without_CS_Courses, useNA = "ifany")

#Hence, we can export the final clean data
write.csv(
  combined_data_clean1,
  file = "output/combined_data_clean_final.csv",
  row.names = FALSE
)
```

5. **Exploratory Visualizations**:  
   - Excerpts from the data
   - Data Dictionary  
```{r, echo = FALSE}
##A sample of the data set and full data structure
head(combined_data_clean1, 5)
str(combined_data_clean1)


##Here is the data dictionary–style table, showing, One row per variable, Column for variable name, Column for type (numeric, character, factor, etc.), Column for levels if the variable has levels (like factors or character variables with few unique values). This will be very useful for checking your dataset and making summary tables.
# Create a data dictionary table
data_dict <- tibble(
  Variable = names(combined_data_clean1),
  Type = sapply(combined_data_clean1, class),
  Levels = sapply(combined_data_clean1, function(x) {
    if (is.factor(x) || is.character(x)) {
      n_levels <- length(unique(na.omit(x)))
      if (n_levels <= 10) {
        paste(sort(unique(na.omit(x))), collapse = ", ")
      } else {
        "Many levels"
      }
    } else {
      NA
    }
  })
)

data_dict

# Export the data dictionary
write.csv(
  data_dict,
  file = "output/data_dictionary.csv",
  row.names = FALSE
)

```


